---
title: "Homework Tidyverse"
author: "Ananya Rajan"
date: "October 6, 2026"
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.height = 3,
  fig.width = 6,
  fig.align = "center"
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
#| warning: false

pacman::p_load(tidyverse, gapminder, broom)
```

# Problem 1:

Copy paste and run the tribble given below.

```{r}
tribble( ~x,    ~y,    ~w,    ~z,
              210,   300,   220,   180,
              102,   100,   119,   187,
              176,   175,   188,   173,
              87,    95,   91,     94,
              202,   210,  234,    218,
              110,   122,  131,    128,
) -> dt
```

## 1_a:

-   Use and show a `map` function to find the "mean" of each column of the `dt` data table

```{r}
map(dt, mean)
```

## 1_b:

-   Use and show a `map` function to find the "standard deviation" of each column of the `dt` data table.

```{r}
map(dt, sd)
```

## 1_c:

-   Use and show a map function that will calculate the "square root" of each value of each column of the data table `dt`.

```{r}
map(dt, sqrt)
```

## 1_d:

-   Use R code to find the "mean", "max", "1st Quartile", "3rd Quartile", "Median", and "Min" for each column of the `dt` data table. (Hint: You do not have to use a map function)

```{r}
summary(dt)
```

# Problem 2

## 2_a:

Repeat part 1 by using a `map_*()` function.

```{r}
means <- map_dbl(1:10000, ~mean(rexp(5, rate = 1)))

hist(means,
     main = "Distribution of Sample Means",
     xlab = "Mean",
     breaks = 30)
```

## 2_b:

Repeat part 1 by using the `replicate()` function.

```{r}
means <- replicate(10000, mean(rexp(5, rate = 1)))

hist(means,
     main = "Distribution of Sample Means",
     xlab = "Mean",
     breaks = 30)
```

## 2_c:

Use another for loop that will print out plots for sample sizes of 5, 10, and 20 observations.

```{r}
sample_sizes <- c(5, 10, 20)

for (n in sample_sizes) {
  
  means <- replicate(10000, mean(rexp(n, rate = 1)))
  
  hist(means,
       main = paste("Sample Size =", n),
       xlab = "Mean",
       breaks = 30)
}
```

# Problem 3:

Use and show R coding to calculate the "standard deviation" for each variable of the data table `mtcars` using the "Special For Loop Method".

```{r}
output <- vector("double", ncol(mtcars))

for (i in seq_along(mtcars)) {
  output[[i]] <- sd(mtcars[[i]])
}

output
```

```{r}
output <- vector("double", ncol(mtcars))

for (i in seq_along(mtcars)) {
  output[[i]] <- sd(mtcars[[i]])
}

names(output) <- names(mtcars)
output
```

# Problem 4

Generate a sequence of dates from January 1, 2015 to December 31, 2025, spaced by every two months. Extract the year, quarter, and ISO week number for each date.

```{r}
dates <- seq.Date(
  from = as.Date("2015-01-01"),
  to = as.Date("2025-12-31"),
  by = "2 months"
)

problem4 <- data.frame(
  date = dates,
  year = year(dates),
  quarter = quarter(dates),
  iso_week = isoweek(dates)
)

problem4
```

# Problem 5

Given the following dates, compute the difference in months and weeks between each consecutive pair.

sample_dates \<- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")

```{r}
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")

sample_dates <- as.Date(sample_dates)

months_diff <- interval(sample_dates[-length(sample_dates)],
                        sample_dates[-1]) %/% months(1)

weeks_diff <- as.numeric(
  difftime(sample_dates[-1], sample_dates[-length(sample_dates)], units = "weeks")
)

data.frame(
  from = sample_dates[-length(sample_dates)],
  to = sample_dates[-1],
  months = months_diff,
  weeks = weeks_diff
)
```

